---
title: "Art with R"
author: "Me"
date: "`r Sys.Date()`"
output: html_document
---

# Making pictures with base R and plotrix

```{r message=FALSE}
library(svglite)
library(plotrix)
library(ambient)
```

### map range helper

A useful helper function below; it will "map" a number to a new range.

For example, I have test scores out of 80 points but I'd like to have them out of 100.  Someone scored a 50 and I'd like to know their score out of 100.  This helper would let you do: `map(50, 0, 80, 0, 100)`

My example ofc could be done as `100 * 50 / 80`.  Trust me that this helper is pretty useful for cases when you can't easily do mental math to convert ranges.

```{r}
map <- function(x, x_lo, x_hi, new_lo, new_hi) {
  scld <- (x - x_lo) / (x_hi - x_lo)
  scld * (new_hi - new_lo) + new_lo
}

# map 5 from 0 to 10 range to be in 0 to 100 range
map(5, 0, 10, 0, 100) # -> 50

# map 10 from 0 to 10 range to be in 0 to 100 range
map(10, 0, 10, 0, 100) # -> 100

# map 100 from 0 to 200 range to be in 0 to 1 range
map(100, 0, 200, 0, 1) # -> 0.5
```

## Math based visuals

With a helper out of the way.  Let's start with a cosine wave... cause... why not?  If you feel good about making simple plots with R, then this is pretty familiar.

Adam examples based on cosine/sine waves:
* https://www.instagram.com/p/CRWiSYioNXH/
* https://www.instagram.com/p/CVVc6dXlkD6/
* https://www.instagram.com/p/CWTXCc-F8X9/

```{r}
x <- 1:(100 * pi) / 20
y <- cos(x)

plot(x, y, type = "l")
```

### Removing chart elements

Axes are of course needed if you're trying to do analytics work, but we're trying to make pretty pictures!

Let's do away with the pesky axes so we can have a prettier picture!

```{r}
plot(
  x, y,
  type = "l", # plot a line instead of points
  xlab = "", # no x axis label
  ylab = "", # no y axis label
  axes = FALSE, # turn off axes
  frame.plot = FALSE # turn of the outline around the plot
)
```

That's a lot of code to keep track of.  This is the perfect use case for a function.

Convert the above to a function named `plot_lines_only`.  This function should accept 2 arguments: `x` and `y`.

```{r}

```

Prove it produces the same output:

```{r}
plot_lines_only(x, y)
```

Just one line is boring.... let's add another wave with smaller peaks.  We can use the `lines()` function to add to an existing plot.

```{r}
plot_lines_only(x, y)
lines(x, y * 0.5)
```

Add programming to add flair! Let's add a bunch of waves at decreasing heights.

Trivia question! What's the height of a wave called?

* Loop through the values 1 to 10
* Use `map()` to change the 1-10 number to be in range 0.9-0.1
* Plot the original x, y using `lines()` but multiply y by this new number we've created

```{r}
# initial plot
plot_lines_only(x, y)

# loop through values to scale y by
for () {
  # convert my i from its 1-10 range
  # get a number back in range 1 to 0.1
  scl <- 
  
  # Use lines to plot
  lines(x, y * scl)
}
```

* Play with things!
  * change the number of waves being made
  * play with the values in map
  * try adding a little bit to x each time (ie x + i / 30)

```{r}
plot_lines_only(x, y)

for (i in 1:30) {
  scl <- map(i, 1, 30, 1, 0.1)
  lines(x, y * scl)
}
```

We can even color this pretty easily:

```{r}
color_palette <- rainbow(10)
plot_lines_only(x, y)

for (i in 1:10) {
  scl <- map(i, 1, 10, 1, 0.1)
  lines(x, y * scl, col = color_palette[i])
}
```

Other ideas to play with:

* Does this look cooler with sin?
* Copy paste the loop (2 loops->2 sets of waves) and play with the values in map
* Overlay sin and cos
* Overlay this wave with a flipped version of itself

### Saving visuals

Is this art yet? Trick question ofc if you think it's art, it's art. If you want to save it there's a couple options.

Save as a png:
```r
png('my_aRt.png')

# your plot code goes here

dev.off()
```

Save as an svg (this is what my pen plotter needs):
```r
library(svglite)

svglite('my_aRt.svg')

# your plot code goes here

dev.off()
```

Save our work so far:

```{r}
png("my_aRt.png")

plot_lines_only(x, y)

for (i in 1:10) {
  scl <- map(i, 1, 10, 1, 0.1)
  lines(x, y * scl)
}

dev.off()
```

## Drawing shapes

### Blank canvas helper

A helper to create a plot canvas with equal axis lengths. 

```{r}
# A helper to start a blank canvas
# Defaults to a range from 1 to 100
blank_canvas <- function(rng = 1:100) {
  plot(
    rng,
    asp = 1,
    type = "n", xlab = "", ylab = "",
    axes = FALSE, frame.plot = FALSE
  )
}

# Very boring output (should be blank)
blank_canvas()
```

### Basic shapes

* Use rect and draw.circle to place shapes on the canvas

```{r}
blank_canvas(1:100)

rect(xleft = 0, ybottom = 0, xright = 50, ytop = 50)
draw.circle(x = 50, y =  50, radius = 10)
```

There might be a triangle function somewhere... or you can make your own!

```{r}
triangle <- function(x1, y1, x2, y2, x3, y3, border = "black", ...) {
  lines(c(x1, x2), c(y1, y2), col = border, ...)
  lines(c(x2, x3), c(y2, y3), col = border, ...)
  lines(c(x3, x1), c(y3, y1), col = border, ...)
}


blank_canvas(1:100)

rect(xleft = 0, ybottom = 0, xright = 50, ytop = 50)
draw.circle(25, 50, 20)
triangle(20, 20, 45, 65, 70, 20)
```

```{r}
blank_canvas(1:100)

rect(0, 0, 50, 50, border = "red")
draw.circle(25, 50, 20, border = "green")
triangle(20, 20, 45, 65, 70, 20, border = "blue")
```

### Making a grid of shapes

First we can make a row of shapes:

```{r}
blank_canvas(1:100)

radius <- 4
y <- 50

ncols <- 10
for (i in 1:ncols) {
  x <- map(i, 1, ncols, 1, 100)
  draw.circle(x, y, radius)
}
```

Extending this loop idea, instead of just plotting a single row.  We can loop over different values of y and create rows at different heights. With the below nested loop we can display a grid of circles.

```{r}
blank_canvas(1:100)

# Change values of radius, nrows, and ncols
#   to see different patterns
radius <- 4

nrows <- 10
ncols <- 10
for (i in 1:nrows) {
  y <- map(i, 1, nrows, 1, 100)
  # row code from before (could make this a function)
  for (j in 1:ncols) {
    x <- map(j, 1, ncols, 1, 100)
    draw.circle(x, y, radius)
  }
}
```

With this circle grid framework you can make some interesting looking patterns by using different values of radius.

Instead of setting a radius for every circle, we can base the radius on the location of the circle.  Try out things like:

* x + y, x * y, etc
* make up your own!

```{r}
blank_canvas(1:100)

nrows <- 10
ncols <- 10
for (i in 1:nrows) {
  y <- map(i, 1, nrows, 1, 100)
  for (j in 1:ncols) {
    x <- map(j, 1, ncols, 1, 100)

    # change this! find something cool
    radius <- 10
    draw.circle(x, y, radius)
  }
}
```

Here's an example finding the distance between the grid point and the grid center.  Then we can "map" that distance between a min and a max value of radius (farther from center -> larger radius)

What else could you map radius to?

```{r}
pt_dist <- function(x1, y1, x2, y2) {
  sqrt((x1 - x2)^2 + (y1 - y2)^2)
}

blank_canvas(1:100)

# center of grid is at 50, 50
cx <- 50
cy <- 50

min_radius <- 0.1
max_radius <- 10

nrows <- 25
ncols <- 25
for (i in 1:nrows) {
  y <- map(i, 1, nrows, 1, 100)
  for (j in 1:ncols) {
    x <- map(j, 1, ncols, 1, 100)

    dist_to_center <- pt_dist(x, y, cx, cy)
    radius <- map(dist_to_center, 0, 100, min_radius, max_radius)

    draw.circle(x, y, radius)
  }
}
```

## Noise

One way to end up with some nice organic feeling patterns is to use "noise".  In particular, I've used "Perlin noise" a good bit. With this noise generation process we get a controlled randomness that can be used for generating random terrain in a video game, or.... can be used to size circles in a grid.

Adam examples:
* https://www.instagram.com/p/CX_e2eXl6df/
* https://www.instagram.com/p/CW8jr93liab/
* https://www.instagram.com/p/CVdNAB9FEVA/
* https://www.instagram.com/p/CQyjUI_rNZm/

```{r}
blank_canvas(1:100)

# Play with all of these numbers (esp freq imo)
min_radius <- 0.001
max_radius <- 7

nrows <- 25
ncols <- 25

freq <- 0.01
my_noise <- noise_perlin(c(nrows, ncols), frequency = freq)

for (i in 1:nrows) {
  y <- map(i, 1, nrows, 1, 100)
  for (j in 1:ncols) {
    x <- map(j, 1, ncols, 1, 100)

    radius <- map(
      my_noise[i, j], min(my_noise), max(my_noise), min_radius, max_radius
    )

    draw.circle(x, y, radius)
  }
}
```

More circles!!

* Write a for loop and use `map()` to make a bunch of "concentric" circles

```{r}
blank_canvas()

x <- 50
y <- 50

min_radius <- 2
max_radius <- 20
n_rings <- 10

# Do loop stuff here:
```

Convert the code to be a function named `concentric_circles`.  It should have arguments of:

* x
* y
* max_r
* min_r
* n_rings
* we're gonna also add a fancy `...` thingamajig

```{r}
concentric_circles <- function() {
  
}

blank_canvas()
concentric_circles(50, 50, border = "red")
```

Let's look at what I think is a cool effect known as Moire.

> "When a grid's misaligned with another behind that's a moire!"

quote source: https://xkcd.com/1814/

Examples (just image google of "moire"): https://www.google.com/search?q=moire&tbm=isch

The idea in it's simplest form:
* draw something
* draw the same thing again over the top of the first, but misalign it a little bit

There are other ways to play with this effect.  But this is the above steps are the simplest to express

Adam examples:
* https://www.instagram.com/p/CUu1_41FGDo/ - this ones framed in my office!
* https://www.instagram.com/p/CSKGBNGrNsZ/ - moire from a sine wave!
* https://www.instagram.com/p/CUfY6DVlWbY/
* https://www.instagram.com/p/CZXBIhLl5e8/

Instead of a single set of concentric circles if we overlap them we will see moire patterns appear.  Different ways of overlapping lead to different effects.

* Draw concentric circles 
* Draw concentric circles again but move the center point

```{r}

```

* Do it again and maybe add a 3rd?
* Color the circles?

```{r}

```


We could also put these back in a grid i guess...?

Below is same grid code, but I changed out the circle function to be ours

```{r}
blank_canvas(1:100)

nrows <- 10
ncols <- 10
for (i in 1:nrows) {
  y <- map(i, 1, nrows, 1, 100)
  for (j in 1:ncols) {
    x <- map(j, 1, ncols, 1, 100)
    max_r <- map(x * y, 0, 100**2, 1, 10)

    concentric_circles(x, y, min_r = 0.1, max_r = max_r, n_rings = 5)
  }
}
```

We could then plug the radius back up to noise i guess...?

```{r}
blank_canvas(1:100)

min_radius <- 0.01
max_radius <- 8
n_rings <- 3

nrows <- 25
ncols <- 25

freq <- 0.005
my_noise <- noise_perlin(c(nrows, ncols), frequency = freq)

for (i in 1:nrows) {
  y <- map(i, 1, nrows, 1, 100)
  for (j in 1:ncols) {
    x <- map(j, 1, ncols, 1, 100)
    max_r <- map(
      my_noise[i, j], min(my_noise), max(my_noise), min_radius, max_radius
    )

    concentric_circles(x, y, min_r = 0.1, max_r = max_r, n_rings = n_rings)
  }
}
```

There you have it, a quick whirlwind of different ideas. At the core, try things and see what you like.  Don't take it too seriously and have some fun.  It can be a really thought out thing or just mindlessly tweaking things and hoping something cool appears.

* Make something up!

```{r}

```
